Predicting computational reproducibility of scientific pipelines using collaborative filtering

### 2018-05-07 paper.tex
Uppdated the ALS-Failure figures

% Explain all sampling techniques (maybe rename 4, 5, and 6)
% 1. random unreal (baseline)
\paragraph{Random Unreal}
The training set is sampled randomly with no restriction, that is,
regardless of the file generation time. This method will be used as a
baseline for comparison with other methods, although it is not
realistic. Figure~\ref{fig:Random-Unreal-Sample-Training-set} represents 
a training set sampled using this method.

@@

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution with parameters \textit{\textbf{a}}, \textit{\textbf{c}} and \textit{\textbf{b}} where \textit{\textbf{b}} is set to $N_{f}$ and the other two parameters could be set according to two different approaches. By considering $\alpha$ as training ratio, the distribution average equals to $\alpha N_{f}$ which that means f  $\alpha$ less than $\nicefrac{1}{3}$ the lower limmit \textit{\textbf{a}} would get a negative value in this regard, therefore to avoid this issue the uniform distribution is applied for $\alpha < \nicefrac{1}{3}$. But for $\alpha \leq \nicefrac{1}{3}$ the following approaches are applicable. This distribution aims at sampling files produced towards the end of the pipeline execution.
Two approches are tried to apply this method:
\begin{enumerate}
\item Largest-a (RFNT-L):  
 This approach of triangular distribution with parameters \textit{\textbf{a}}, \textit{\textbf{c}} and \textit{\textbf{b}} where \textit{\textbf{b}} is set to $N_{f}$ and the two other parameters are reperesent the limmited bond and mode of the distribution respectively. The \textit{\textbf{a}} equals to $\nicefrac {N_{f}(3\alpha-1)}{2}$ while \textit{\textbf{c}} is considered as equal as \textit{\textbf{a}}, $(\textit{\textbf{c}} = \textit{\textbf{a}})$.\\

In this approach there is a concern to eliminate the possibility of having less 
than the two first generated files of any subjects in the sampled training set. 
The possibility of missing some subjects in training set can happen to some of them due to 
random selection of subjects or even meeting the determined training ratio before 
having the chance of being selected.

@@

\section{Results}
TP: The number of cases correctly predicted as occurance of difference \\
FP: The number of cases incorrectly predicted as occurance of difference \\
TN: The number of cases correctly predicted as no occurance of difference \\ 
FN: The number of cases incorrectly predicted as no occurance of difference \\
\[
	    Accuracy = \frac{TP+TN}{TP+TN+FP+FN}
\]
Sensitivity: The ability to predict the differences cases correctly. The bellow formula can be used to calculate sensitivity of a test.
\[
            Sensitivity = \frac{TP}{TP+FN}
\]
Specificity: The ability to predict the non differences cases correctly. The following formula can be used to calculate specificity of a test. 
\[
            Specificity = \frac{TN}{TN+FP}
\]

The experimental results have been obtained from four different approaches:\\
1. Applying ALS technique (ALS)\\
2. Employing Bias as well as ALS technique (ALS-Bias)\\
3. Using just Bias (Bias)\\
4. Results stemmed from ommiting the impact of predicted value of canstant files \\
In all above approaches, dummy classifier which makes predictions using simple rules\cite{URL}has been used to indicate that how much predicted value would tend to be anticipate as the dominent value of the utility matrix. This classifier is considered as an accuracy evaluator since the binary utility matrix is consist of \{1,0\} which respectively represent whether a difference has been observed between the same files in two different conditions (by comapring their checksums) or not.
With the knowledge of the fact that Random Unreal method is not applicable into our case study we used its results as the comparision between the unrealistic method and realistics ones. In this regard, the aformentioned approaches are propossed and tried to lessen the gaps between the best results from random-unreal and the applied one.

\subsection{ALS without biases}
In this phase of the experiment, the Alternating Least Squares techniq of Collaborative Filtering is applied for all provided methods. \todo{should I talk about ALS in here?}
\subsubsection{Random Unreal}
Figure~\ref{fig:unreal method} represents
The accuracy graph of Random Unreal method with comparision to dummy
classifier over the increasement of the training ratio.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/unreal-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/unreal-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/unreal-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of Random Unreal methods in three different condition pairs}
        \label{fig:unreal method}
\end{figure}
This method has a great performance with absolute accouracy of more than $0.95$ even by accessing to small taining size of 10 percent. As it mentioned earlier this method is not applicable in pipline due to timestamp cosideration but it is a good criteria to estmiate the performance of other methods and try to modify the proposed methods in order to their accuracy be as colsed as possible to the Random Unreal method ones.

\subsubsection{Complete Columns}

Figure~\ref{fig:columns method} represents
the accuracy graph of Complete Columns method with comparision to dummy
classifier over the increasement of the training ratio.
As it can be seen Complete Columns method has
almost a constant accuracy even if the training ratio increased and it
is almost two times less than dummy classifier in more differentiated
condition pairs (CentOS5 vs CentOS7 and CentOS6 vs CentOS7). But in the
best case which it means in similar conditions the Complete Columns
accuracy is as much accurate as the Dummy classifier which is $0.65$
percent.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/columns-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/columns-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/columns-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of Complete Columns methods in three different condition pairs}
        \label{fig:columns method}
\end{figure}

\subsubsection{Complete Rows}
Figure~\ref{fig:rows method} represents
the accuracy graph of Complete Rows method with comparision to dummy
classifier over the increasement of the training ratio.
As it can be observed, This method has some flactuation over the gowth of training set size. These falactuations are more frequent in condition pairs which have higher difference positive ratio. Also by enlarging the training ratio specially between $0.3$ to $0.7$, we can see dummy classifier has better results than Complete Rows ones. In other words despite of better performance of Complete Rows than dummy classifier at the last 10 percent of the graph, its accuracies are less than dummy classifiers ones.  
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/rows-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/rows-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/rows-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of Complete Rows methods in three different condition pairs}
        \label{fig:rows method}
\end{figure}
\subsubsection{Random Subjects-RS}
Figure~\ref{fig:RS method} represents
the accuracy graph of RS method with comparision to dummy
classifier over the increasement of the training ratio.
Same to the Complete Row method, RS method has some flactuation but not as frequent as Complete Rows method. Althogh there are some up and down trends specialy with training ratio range of $0.2$ to $0.6$ but its constant growth trend after training ratio $0.3$ in condition pairs with high difference positive ratio is considerable. The method can be $0.99$ percent accurate when it has over 70 percent of the data however this accuracy can be achived with higher training ratio (over 80 percent) for C5C6.  
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RS-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RS-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RS-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of Complete RS methods in three different condition pairs}
        \label{fig:RS method}
\end{figure}
\subsubsection{RandomFile Numbers(Uniform) - RFNU}
Figure~\ref{fig:RFNU method} represents
the accuracy graph of RFNU method with comparision to dummy
classifier over the increasement of the training ratio.
The graphs show a continuous rise in the accuracy trend in all condition pairs however a 0.03 drop is recorder in C6C7 beteen training ratio size of $0.8$ to $0.9$. This metod takes over the dummy accuracy after having a training size over 40 percent in C5vsC7 and C6vsC7 but it always shows better acuracy than dummy classifier ($0.73$ percent with traininfg ratio of $0.1$)    
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNU-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNU-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNU-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of RFNU methods in three different condition pairs}
        \label{fig:RFNU method}
\end{figure}
\subsubsection{RNFT-L}
Figure~\ref{fig:RFNT-L method} represents
the accuracy graph of RFNT-L method with comparision to dummy
classifier over the increasement of the training ratio.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-L-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-L-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-L-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of RFNT-L methods in three different condition pairs}
        \label{fig:RFNT-L method}
\end{figure}
\subsubsection{RFNT-S}
Figure~\ref{fig:RFNT-S method} represents
the accuracy graph of RFNTS method with comparision to dummy
classifier over the increasement of the training ratio.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-S-ALS-5vs6-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-S-ALS-5vs7-PFS}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/RFNT-S-ALS-6vs7-PFS}
        \end{subfigure}
        \caption{Average performance of RFNT-S methods in three different condition pairs}
        \label{fig:RFNT-S method}
\end{figure}
% General conclution on results of ALS without Bias 
In general the accuracy pattern of the two Complete Rows and 
RS methods show similar decline and grow behavior over the 
training ratio axis however it can be subject to considerable 
fluctuations in differentiated condition pairs. Dispite of their 
likely irrational behaviour, they always have better accuracy 
for the training size with less than $0.3$ in comparison to other 
methods. This better performance (around 75 percent)is almost three 
times higher than other methods in condition pairs with more 
differences but just 20 percent better when the conditions are more 
similar to each other (CentOS5 vs CentOS6) with accuracy of $90\%$.
Figure~\ref{fig:RS_Rows} represents
the accuracy graph of these two methods with comparision to dummy
classifier over the increasement of the training ratio.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_5vs6}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_5vs7}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_6vs7}
        \end{subfigure}
        \caption{Average performance of RS and Complete Rows methods in three different condition pairs}
        \label{fig:RS_RowS}
\end{figure}

@@

% Summary of the results and discussion. The take-home message.

\section*{Acknowledgment}

>>>>

% 0. random unreal (baseline)
\paragraph{Random Unreal}
The training set is sampled randomly with no restriction, that is,
regardless of the file generation time. This method will be used as a
baseline for comparison with other methods, although it is not
realistic. Figure~\ref{fig:Random-Unreal-Sample-Training-set} represents 
a training set sampled using this method.

@@

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution with parameters \textit{\textbf{a}}, \textit{\textbf{c}} 
and \textit{\textbf{b}} where \textit{\textbf{b}} is set to $N_{f}$ and the other 
two parameters could be set according to two different approaches. By considering $\alpha$ 
as training ratio, the distribution average equals to $\alpha N_{f}$ which that means 
for $\alpha$ less than $\nicefrac{1}{3}$ the lower limmit \textit{\textbf{a}} would 
get a negative value in this regard, therefore to avoid this issue the uniform distribution 
is applied for $\alpha < \nicefrac{1}{3}$. But for $\alpha \leq \nicefrac{1}{3}$ the following 
approaches are applicable. This distribution aims at sampling files produced towards 
the end of the pipeline execution.

Two approches are tried to apply this method:
\begin{enumerate}
\item Largest-a (RFNT-L):  
This approach of triangular distribution with parameters \textit{\textbf{a}}, \textit{\textbf{c}} 
and \textit{\textbf{b}} where \textit{\textbf{b}} is set to $N_{f}$ and the two other 
parameters are reperesent the limmited bond and mode of the distribution respectively. 
The \textit{\textbf{a}} equals to $\nicefrac {N_{f}(3\alpha-1)}{2}$ while \textit{\textbf{c}} 
is considered as equal as \textit{\textbf{a}}, $(\textit{\textbf{c}} = \textit{\textbf{a}})$.\\

In this approach there is a concern to eliminate the possibility of having less 
than the first two generated files of any subjects in the sampled training set. 
The possibility of missing some subjects in training set can happen to some of them due to 
random selection of subjects or even meeting the determined training ratio before 
having the chance of being selected.

@@

\section{Results}

TP: The number of cases correctly predicted as occurance of difference \\
FP: The number of cases incorrectly predicted as occurance of difference \\
TN: The number of cases correctly predicted as no occurance of difference \\ 
FN: The number of cases incorrectly predicted as no occurance of difference \\
\[
	    Accuracy = \frac{TP+TN}{TP+TN+FP+FN}
\]
Sensitivity: The ability to predict the differences cases correctly. The bellow formula can be used to calculate sensitivity of a test.
\[
            Sensitivity = \frac{TP}{TP+FN}
\]
Specificity: The ability to predict the non differences cases correctly. The following formula can be used to calculate specificity of a test. 
\[
            Specificity = \frac{TN}{TN+FP}
\]

The experimental results have been obtained from four different approaches:\\
1. Applying ALS technique (ALS)\\
2. Employing Bias as well as ALS technique (ALS-Bias)\\
3. Using just Bias (Bias)\\
4. Results stemmed from ommiting the impact of predicted value of canstant files \\
In all above approaches, dummy classifier which makes prediction by using simple rules\cite{URL}
has been used to indicate that how much the predicted value would tend to be anticipate 
as the dominent value of the utility matrix. This classifier is considered as an 
accuracy evaluator since the binary utility matrix is consist of \{1,0\} which  
respectively represent whether a difference has been observed between the same files 
in two different conditions (by comapring their checksums) or not.
With the knowledge of the fact that Random Unreal method is not applicable into our 
case study, we used its results as the comparision between the unrealistic method 
and realistics ones. In this regard, the aformentioned approaches are propossed and 
tried to lessen the gaps between the best results from random-unreal and the applied one.

\subsection{ALS without biases}
In this phase of the experiment, the Alternating Least Squares techniq of Collaborative Filtering is applied for all provided methods. \todo{should I talk about ALS in here?}
Figure~\ref{fig:ALS-Failure} represents
The accuracy graph of four diferent Complete Columns, RS, Complete Rows and Random Unreal methods.The comparision to dummy
classifier over the increasement of the training ratio can be seen in this figure.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/AlS-Failure-5vs6}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/AlS-Failure-5vs7}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/AlS-Failure-6vs7}
        \end{subfigure}
        \caption{Average performance of Complete Columns, RS, Complete Rows and Random Unreal methods in three different condition pairs}
        \label{fig:ALS-Failure}
\end{figure}

\subsubsection{Random Unreal}
This method has a great performance with absolute accouracy of more than $0.95$ 
even by accessing to small taining size of 10 percent. As it mentioned earlier 
this method is not applicable in pipline due to timestamp cosideration but it 
is a good criteria to estmiate the performance of other methods and try to modify 
the proposed methods in order to their accuracy be as colsed as possible to the 
Random Unreal method ones.

\subsubsection{Complete Columns}
As it shown in Figure~\ref{fig:ALS-Failure} the Complete Columns method has
almost a constant accuracy even if the training ratio increased and it
is almost two times less than dummy classifier in more differentiated
condition pairs (CentOS5 vs CentOS7 and CentOS6 vs CentOS7). But in the
best case which it means in similar conditions the Complete Columns
accuracy is as much accurate as the Dummy classifier which is $0.65$
percent.
\\
\subsubsection{Complete Rows}
As it can be observed, Figure~\ref{fig:ALS-Failure}, This method has some 
flactuation over the gowth of training set size. These falactuations are more 
frequent in condition pairs which have higher difference positive ratio. 
Also by enlarging the training ratio specially between $0.3$ to $0.7$, 
we can see dummy classifier has better results than Complete Rows ones. 
In other words despite of better performance of Complete Rows than 
dummy classifier at the last 10 percent of the graph, its accuracies 
are less than dummy classifiers ones.  
\\
\subsubsection{Random Subjects-RS}
The performance of RS method is represented in Figure~\ref{fig:ALS-Failure}.
Same to the Complete Row method, RS method has some flactuation but not as frequent 
as Complete Rows method. Althogh there are some up and down trends specialy with 
training ratio range of $0.2$ to $0.6$ but its constant growth trend after 
training ratio $0.3$ in condition pairs with high difference positive ratio is considerable.
The method can be $0.99$ percent accurate when it has over 70 percent of the data however 
this accuracy can be achived with higher training ratio (over 80 percent) for C5C6.  
\\
\\
\noindent Figure~\ref{fig:ALS-Succeed} represents
the accuracy graph of all Random kind methods with comparision to dummy
classifier over the increasement of the training ratio.\\
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/ALS-Succeed-5vs6}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/ALS-Succeed-5vs7}
        \end{subfigure}
        \begin{subfigure}[b]{0.8\linewidth}
                \includegraphics[width=\columnwidth]{figures/ALS/ALS-Succeed-6vs7}
        \end{subfigure}
        \caption{Average performance of RFNU methods in three different condition pairs}
        \label{fig:ALS-Succeed}
\end{figure}
\subsubsection{Random File Numbers(Uniform) - RFNU}
The continuous rise of RFNU accuracy trend can be observed in Figure~\ref{fig:ALS-Succeed}.
However a 0.03 drop is recorded in C6C7 between training ratio size of $0.8$ to $0.9$. 
This method takes over the dummy accuracy after having a training size more than 40 percent 
in condition pairs of C5vsC7 and C6vsC7 but it always shows better acuracy than 
dummy classifier ($0.73$ percent with traininfg ratio of $0.1$).

As it mentioned earlier, for $\alpha < \nicefrac{1}{3}$ RFNT approaches are not applicable, 
therefor for training ratio less than $0.4$ RFNU method is employed. As a result, in terms of 
evaluating the performance of RFNT-L and RFNT-S approches, only those accuracy results are 
considered which their training set are equal or geather than 0.4 percent of the whole utility matrix.

\subsubsection{RFNT-L}
Figure~\ref{fig:ALS-Succeed}

\subsubsection{RFNT-S}
Figure~\ref{fig:ALS-Succeed}

% General conclution on results of ALS without Bias 
In general the accuracy pattern of the two Complete Rows and 
RS methods show similar decline and grow behavior over the 
training ratio axis however it can be subject to considerable 
fluctuations in differentiated condition pairs. Dispite of their 
likely irrational behaviour, they always have better accuracy 
for the training size with less than $0.3$ in comparison to other 
methods. This better performance (around 75 percent)is almost three 
times higher than other methods in condition pairs with more 
differences but just 20 percent better when the conditions are more 
similar to each other (CentOS5 vs CentOS6) with accuracy of $90\%$.
Figure~\ref{fig:RS_Rows} represents
the accuracy graph of these two methods with comparision to dummy
classifier over the increasement of the training ratio.
\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_5vs6}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_5vs7}
        \end{subfigure}
        \begin{subfigure}[b]{0.4\linewidth}
                \includegraphics[width=\columnwidth]{figures/RS_Rows_6vs7}
        \end{subfigure}
        \caption{Average performance of RS and Complete Rows methods in three different condition pairs}
        \label{fig:RS_RowS}
\end{figure}

@@

% Summary of the results and discussion. The take-home message.
%%%%%% why there is a drop in ALS-RFNT-L with 60% ratio?  
\section*{Acknowledgment}
